Active Learning for Level Set Estimation

### 2013-03-20 msthpres.tex
LSE theory and experiments

\documentclass[10pt]{beamer}

@@

\setbeamertemplate{section in toc}{$\circ$ \inserttocsection}
\setbeamertemplate{itemize item}{\scriptsize\raise1.25pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{itemize subitem}{\tiny\raise1.5pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{itemize subsubitem}{\tiny\raise1.5pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{enumerate item}{\insertenumlabel.}
\setbeamertemplate{enumerate subitem}{\insertenumlabel.\insertsubenumlabel}
\setbeamertemplate{enumerate subsubitem}{\insertenumlabel.\insertsubenumlabel.\insertsubsubenumlabel}
\setbeamertemplate{enumerate mini template}{\insertenumlabel}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{libertine}
\usepackage[T1]{fontenc}

%%
\usepackage{xspace}
\usepackage{msthpres}
\newcommand{\gpucb}{\textsf{GP-UCB}\xspace}
\newcommand{\gpbucb}{\textsf{GP-BUCB}\xspace}
\newcommand{\acl}{\textsf{LSE}\xspace}
\newcommand{\fullacl}{Level Set Estimation\xspace}
\newcommand{\iacl}{\textsf{LSE\textsubscript{imp}}\xspace}
\newcommand{\bacl}{\textsf{LSE\textsubscript{batch}}\xspace}
\newcommand{\ibacl}{\textsf{LSE\textsubscript{imp-batch}}\xspace}
\newcommand{\str}{\textsf{STR}\xspace}
\newcommand{\istr}{\textsf{STR\textsubscript{imp}}\xspace}
\newcommand{\bstr}{\textsf{STR\textsubscript{batch}}\xspace}
\newcommand{\rstr}{\textsf{STR\textsubscript{rank}}\xspace}
\newcommand{\var}{\textsf{VAR}\xspace}
%%

@@

\section{Motivation and Background}

@@

\begin{frame}
\begin{center}
\uncover<1->{\large Why GPs are nice}
\end{center}
\begin{itemize}
\item<2-> \textbf{Nonparametric}:\\
      Impose ``smoothness'' assumptions via kernel $k(x, x')$.
\vspace{1em}
\item<3-> \textbf{Bayesian}, yet \textbf{efficient}:
      \begin{align*}
        \mu_t(\*x) &= \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*y_t\\
        k_t(\*x, \*x') &= k(\*x, \*x') - \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*k_t(\*x)\notag\\
        \sigma_t^2(\*x) &= k_t(\*x, \*x)
      \end{align*}
      Suitable for step-by-step updates.
\vspace{1em}
\item<4-> More importantly: they provide \textbf{variance estimates}!\\
          Construct confidence intervals: $Q_t(\*x) = \left[\mu_{t-1}(\*x) \pm \beta_t \sigma_{t-1}(\*x)\right]$
\end{itemize}
\end{frame}

@@

\definecolor{c1}{RGB}{255,255,255}
\definecolor{c2}{RGB}{27,161,226}
\definecolor{c6}{RGB}{140,191,38}
\definecolor{c4}{RGB}{240,150,9}
\definecolor{c5}{RGB}{255,255,255}
\definecolor{c3}{RGB}{230,113,184}
%\colorlet{c1}{cyan}
%\colorlet{c2}{magenta}
%\colorlet{c3}{red}
%\colorlet{c4}{green}
%\colorlet{c5}{yellow}

@@

\section{The LSE algorithm}

@@

\section{Results}

@@

%\begin{frame}{Make Titles Informative. Use Uppercase Letters.}{Subtitles are optional.}
%  % - A title should summarize the slide in an understandable fashion
%  %   for anyone how does not follow everything on the slide itself.
%
%  \begin{itemize}
%  \item
%    Use \texttt{itemize} a lot.
%  \item
%    Use very short sentences or short phrases.
%  \end{itemize}
%\end{frame}
%
%\begin{frame}{Make Titles Informative.}
%
%  You can create overlays\dots
%  \begin{itemize}
%  \item using the \texttt{pause} command:
%    \begin{itemize}
%    \item
%      First item.
%      \pause
%    \item    
%      Second item.
%    \end{itemize}
%  \item
%    using overlay specifications:
%    \begin{itemize}
%    \item<3->
%      First item.
%    \item<4->
%      Second item.
%    \end{itemize}
%  \item
%    using the general \texttt{uncover} command:
%    \begin{itemize}
%      \uncover<5->{\item
%        First item.}
%      \uncover<6->{\item
%        Second item.}
%    \end{itemize}
%  \end{itemize}
%\end{frame}

%\begin{frame}{Summary}
%
%  % Keep the summary *very short*.
%  \begin{itemize}
%  \item
%    The \alert{first main message} of your talk in one or two lines.
%  \item
%    The \alert{second main message} of your talk in one or two lines.
%  \item
%    Perhaps a \alert{third message}, but not more than that.
%  \end{itemize}
%  
%  % The following outlook is optional.
%  \vskip0pt plus.5fill
%  \begin{itemize}
%  \item
%    Outlook
%    \begin{itemize}
%    \item
%      Something you haven't solved.
%    \item
%      Something else you haven't solved.
%    \end{itemize}
%  \end{itemize}
%\end{frame}

>>>>

\documentclass[10pt,mathserif,serif]{beamer}
%\documentclass[10pt,mathserif,serif,draft, handout]{beamer}

@@

\setbeamertemplate{section in toc}{$\circ$ \inserttocsection}
\setbeamertemplate{subsection in toc}{\hspace{1em}{\footnotesize$\circ$} \inserttocsubsection}
\setbeamertemplate{itemize item}{\scriptsize\raise1.25pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{itemize subitem}{\tiny\raise1.5pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{itemize subsubitem}{\tiny\raise1.5pt\hbox{\donotcoloroutermaths$\blacktriangleright$}}
\setbeamertemplate{enumerate item}{\insertenumlabel.}
\setbeamertemplate{enumerate subitem}{\insertenumlabel.\insertsubenumlabel}
\setbeamertemplate{enumerate subsubitem}{\insertenumlabel.\insertsubenumlabel.\insertsubsubenumlabel}
\setbeamertemplate{enumerate mini template}{\insertenumlabel}

\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Linux Libertine}
%\setsansfont{Myriad Pro}
%\setmonofont[Scale=0.9]{Consolas}
\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}
%\usepackage{libertine}
%\usepackage[T1]{fontenc}

%%
\usepackage{xspace}
\usepackage{makecell}
\usepackage{msthpres}
\newcommand{\gpucb}{\textsf{GP-UCB}\xspace}
\newcommand{\gpbucb}{\textsf{GP-BUCB}\xspace}
\newcommand{\acl}{\textsf{LSE}\xspace}
\newcommand{\fullacl}{Level Set Estimation\xspace}
\newcommand{\iacl}{\textsf{LSE\textsubscript{imp}}\xspace}
\newcommand{\bacl}{\textsf{LSE\textsubscript{batch}}\xspace}
\newcommand{\ibacl}{\textsf{LSE\textsubscript{imp-batch}}\xspace}
\newcommand{\str}{\textsf{STR}\xspace}
\newcommand{\istr}{\textsf{STR\textsubscript{imp}}\xspace}
\newcommand{\bstr}{\textsf{STR\textsubscript{batch}}\xspace}
\newcommand{\rstr}{\textsf{STR\textsubscript{rank}}\xspace}
\newcommand{\var}{\textsf{VAR}\xspace}
%%

\usepackage{tikz}
\usetikzlibrary{arrows,positioning,shapes,fit}
\tikzset{
    %Define standard arrow tip
    >=stealth',
    %Define style for boxes
    st/.style={
           rectangle,
           %rounded corners,
           draw=black, thick,
           %text width=6.5em,
           minimum height=2em,
           text centered},
    % Define arrow style
    pil/.style={
           ->,
           semithick,
           shorten <=2pt,
           shorten >=2pt,}
}

@@

\section{Motivation and background}

@@

\begin{frame}
\begin{center}
\uncover<1->{\large Why GPs are nice}
\end{center}
\begin{itemize}
\item<2-> \textbf{Nonparametric}:\\
      Impose ``smoothness'' assumptions via kernel $k(x, x')$.
\vspace{1em}
\item<3-> \textbf{Bayesian}, yet \textbf{efficient}:
      \begin{align*}
        \mu_t(\*x) &= \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*y_t\\
        k_t(\*x, \*x') &= k(\*x, \*x') - \*k_t(\*x)^T\left(\*K_t + \sigma^2 \*I\right)^{-1}\*k_t(\*x)\notag\\
        \sigma_t^2(\*x) &= k_t(\*x, \*x)
      \end{align*}
      Suitable for step-by-step updates.
\vspace{1em}
\item<4-> More importantly: they provide \textbf{variance estimates}!\\
          Construct confidence intervals: $Q_t(\*x) = \left[\mu_{t-1}(\*x) \pm \beta_t^{1/2} \sigma_{t-1}(\*x)\right]$
\end{itemize}
\end{frame}

@@

\section{The LSE algorithm}
\definecolor{c1}{RGB}{255,255,255}
\definecolor{c2}{RGB}{27,161,226}
\definecolor{c6}{RGB}{140,191,38}
\definecolor{c4}{RGB}{240,150,9}
\definecolor{c5}{RGB}{255,255,255}
\definecolor{c3}{RGB}{230,113,184}
%\colorlet{c1}{cyan}
%\colorlet{c2}{magenta}
%\colorlet{c3}{red}
%\colorlet{c4}{green}
%\colorlet{c5}{yellow}

@@

\subsection*{Theoretical analysis}
\begin{frame}
How do we judge the ``accuracy'' of a solution $(\hat{H}, \hat{L})$?
\uncover<2->{
\begin{definition}[Misclassification loss]
%\begin{align*}
$\ell_h(\*x) = \twopartdef{\max\{0, f(\*x) - h\}}{\*x\in \hat{L}}{\max\{0, h - f(\*x)\}}{\*x\in \hat{H}}$
%\end{align*}
\end{definition}
\begin{center}
\includegraphics[width=3.5in]{figures/misclass}
\end{center}
}
\end{frame}

\begin{frame}
\footnotesize
\begin{tikzpicture}[node distance=1cm, auto,]
 \uncover<1->{
 \node[st] (conv) {Convergence};
 }

 \uncover<3->{
 \node[st,below=3em of conv] (clas) {All points in $D$ have been classified}
   edge[pil] (conv);
 }

 \uncover<4->{
 \node[st,text width=14em,below=3em of clas] (amb)
 {\makecell[c]
   {Conf. intervals ($\Leftrightarrow$ ambiguities)\\
    get small enough ($<2\epsilon$)\uncover<12->{:\\[0.5em]
    $a_t(\*x_t) \sim \mathcal{O}\left(\sqrt{\displaystyle\frac{\beta_t\gamma_t}{t}}\right)$
    }
   }
 } edge[pil] (clas);

 \uncover<4->{ 
 \node[st,right=1.5em of clas] (eps) {$\epsilon$}
   edge[pil] (amb);
 }

 \uncover<18->{ 
 \node[st,right=1.5em of clas,fill=c6] (eps) {$\epsilon$}
   edge[pil] (amb);
 }

 \node[st,draw=none,below=3em of amb] (dummy) {};
 }

 \uncover<11->{
 \node[st,left=2em of amb] (t0) {Monotonicity}
   edge[pil] (amb);
 }

 \uncover<5->{
 \node[st,left=3em of dummy] (t1) {Samples ($t$)}
   edge[pil] (amb);
 }

 \uncover<6->{
 \node[st,right=1em of t1] (t2) {Kernel ($k$)}
   edge[pil] (amb);
 }

 \uncover<7->{
 \node[st,right=1em of t2] (t3) {Noise ($\sigma$)}
   edge[pil] (amb);
 }

 \uncover<8->{
 \node[st,right=3em of t3] (t4) {$\beta_t$}
   edge[pil] (amb);
 }

 \uncover<18->{
 \node[st,right=3em of t3,fill=c6] (t4) {$\beta_t$}
   edge[pil] (amb);
 }

 \uncover<9->{
 \node[draw=black,densely dotted,line width=0.7pt,fit=(t1)(t3)] (fit1) {};}

 \uncover<9->{
 \node[st,below=3em of fit1] (gamma)
 {\makecell[c]
   {Quantify by max. information gain (Srinivas \emph{et al.}, 2010):\\[0.5em]
    $\gamma_t = \max_{\*y_t} I(\*y_{1:t}; f)$\uncover<10->{\\[0.5em]
    For SE kernel $\gamma_t \sim \mathcal{O}\left((\log t)^{d+1}\right)$
    }
   }
 } edge[pil] (fit1);
 }

 \uncover<2->{
 \node[st,right=9em of conv] (acc) {Solution accuracy};
 }

 \uncover<14->{
 \node[st,below=3em of acc] (loss) {$\max_{\*x\in D} \ell_h(\*x) \leq \epsilon$}
   edge[pil] (acc)
   edge[pil,<-] (eps);
 }

 \uncover<15->{
 \node[st,right=2em of loss] (rules) {Class. rules}
   edge[pil] (loss);
 }

 \uncover<16->{ 
 \node[st,below=3em of loss] (est) {Correctly estimate $f$ w.h.p.}
   edge[pil] (loss);
 }

 \uncover<16->{ 
 \node[st,right=2em of est] (delta) {$\delta$}
   edge[pil] (est);
 }

 \uncover<17->{
 \node[st,below=3em of est] (large) {Conf. intervals are large enough}
   edge[pil] (est)
   edge[pil,<-] (t4);
 }
\end{tikzpicture}
\end{frame}

\begin{frame}
\begin{theorem}[Convergence of \acl]
\label{thm:acl}
For any $h\in\mathbb{R}$, $\delta \in (0, 1)$, and $\epsilon > 0$,
if $\beta_t = 2\log(|D|\pi^2 t^2/(6\delta))$, \acl terminates after
at most $T$ iterations, where $T$ is the smallest positive integer
satisfying
\begin{align*}
\frac{T}{\beta_T \gamma_T} \geq \frac{C_1}{4\epsilon^2},
%T/(\beta_T \gamma_T) \geq C_1/(4\epsilon^2),
\end{align*}
where $C_1 = 8 / \log(1 + \sigma^{-2})$.

Furthermore, with probability at least $1-\delta$, the algorithm returns
an $\epsilon$-accurate solution, that is
\begin{align*}
\Pr\left\{\max_{\*x\in D}\ell_h(\*x) \leq \epsilon\right\} \geq 1 - \delta.
\end{align*}
\end{theorem}
\end{frame}

\subsection*{Results}
\begin{frame}
\begin{center}
{\large Experiments}
\end{center}
\begin{enumerate}
\item \acl
\item<2-> Maximum variance sampling:
\vspace{-0.5em}
\begin{align*}
\*x_t = \argmax_{\*x\in D} \sigma_{t-1}(\*x)
\end{align*}
\item<3-> ``Straddle'' heuristic (Bryan \emph{et al.}, 2005):
\vspace{-0.5em}
\begin{align*}
\*x_t &= \argmax_{\*x\in D} 1.96\sigma_{t-1}(\*x)-|\mu_{t-1}(\*x) - h|\\
&\approx \argmax_{\*x\in D} a_{t-1}(\*x) \tag{for $\beta_t = 1.96$}
\end{align*}
\end{enumerate}
\end{frame}

\begin{frame}
\begin{center}
\hspace{4pt}\includegraphics<1->[width=4in]{figures/limno_chl_ls}\\[0.5em]
\uncover<2->{\includegraphics[width=2.3in]{figures/ev_chl_seq}}
\uncover<3>{\includegraphics[width=2.3in]{figures/ev_chl_eps}}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics<1->[width=4in]{figures/limno_bgape_ls}\\[0.5em]
\color{white}
\only<1>{\includegraphics[draft,width=2.3in]{figures/ev_bgape_seq}}
\color{black}
\only<2-4>{\includegraphics[width=2.3in]{figures/ev_bgape_seq}}
\color{white}
\only<2>{\includegraphics[draft,width=2.3in]{figures/ev_bgape_eps}}
\color{black}
\only<3>{\includegraphics[width=2.3in]{figures/ev_bgape_eps}}
\only<4>{\includegraphics[width=2.3in]{figures/ev_bgape_seq_str_sc}}
\only<5>{\includegraphics[width=4in]{figures/limno_bgape_class_354}}
\only<6>{\includegraphics[width=4in]{figures/limno_bgape_str_class_400}}
\only<7>{\includegraphics[width=4in]{figures/limno_bgape_var_class_400}}
\end{center}
\end{frame}

@@

@@
